\mytitle{}

### 2019-05-14 paper/paper.tex
Post proof-reading

\begin{abstract}
%?? Trim to 75--200 words, simplify.
DDoS attacks plague the availability of online services today, yet like many cybersecurity problems are evolving and non-stationary.
Normal and attack patterns shift as new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
Accordingly, application of machine learning-based techniques is difficult.
%Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.

@@

Our results show the existence of flow features with high predictive power for different traffic classes, when a basis for feedback-loop-like control.
We show that the new RL agent models can offer a significant increase in goodput of legitimate TCP traffic for many choices of host density.

@@

To date, there have been few applications of this class of algorithms towards intrusion detection and prevention which make use of their full potential for online control, rather than using them as the basis for a classifier.
We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
%?? Expand as required
What approaches do exist are aimed towards the task of adaptive online DDoS mitigation, and rely upon learning control of probabilistic packet drop.
%?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An empirical evaluation of our new models alongside the state-of-the-art in RL-based DDoS mitigation with these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
\end{itemize}

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this either by exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.4\linewidth}
		\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
		\node at (0,0.3){$s = \begin{pmatrix}
			0.7 \\
			0.3
			\end{pmatrix}$};
		\node at (0,-1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
			T_{1,9}, \\
			T_{2,5}, \\
			T_{\mathit{bias}}
			\end{Bmatrix}$};

@@

			\draw[step=0.5cm,color=uofgpumpkin,opacity=0.7,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
			\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
			\node[color=uofgpumpkin] at (0.25,-0.95) {\footnotesize Tiling 2};

@@

		\end{tikzpicture}
		}
		\caption{Tile Coding\label{fig:tilecode-code}}
	\end{subfigure}%
	\begin{subfigure}{0.4\linewidth}
		\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
		% Top half
		\def\topacs{
			{-0.3,-0.5,-0.1,0.8},
			{0.1,0.1,-0.2,0.3},
			{0.3,0.4,0.2,-0.4}%
		}

@@

This combination of algorithm and coding strategy is well-optimised, if actions are discrete; this allows a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each tile.
Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
This means that we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update:
\begin{equation}
\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
\label{eqn:sg-sarsa-opt}
\end{equation}
\Cref{fig:tilecode-update} shows how this applies to our prior example.
If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.
A caveat of tile coding remains, in that the value of $\alpha$ must be reduced according to the number of tilings to prevent divergence at the expense of slower learning ($\alpha \leftarrow \alpha / n_{\mathit{tilings}}$).

@@

\subsubsection{Action rate}
We chose to adapt the algorithm to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
To this end, we allow agents to make many decisions per timestep.
We maintain the last state-action pair associated with each source IP and destination, and calculate any actions for the flows which still exist.
Finally, we update $\wvec{}$ using each available trace and the reward signal from the relevant destination.
As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
In turn, we increase the annealing window for $\epsilon$ by a factor of 8 so as to preserve exploration over time, by accounting for the greater volume of decisions being made.

@@

\begin{figure}
	\centering
	\resizebox{0.7\linewidth}{!}{
		\begin{tikzpicture}[
		texts/.style = {text=black},
		labeltexts/.style = {text=gray},
		treeline/.style = {draw=uofgburgundy},
		treenode/.style = {texts, circle, centered, fill=white, treeline},
		load/.style = {fill=uofgcobalt},
		loadline/.style = {draw=uofgcobalt, line width=0.75mm},
		loadhide/.style = {fill=uofgcobalt!40!white},
		external/.style = {fill=uofgrust},
		externalhide/.style = {fill=uofgrust!40!white},
		hideline/.style = {draw=uofgsandstone!40!white},
		hidenode/.style = {treenode, hideline},
		grow'=right
		]
		\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
		\node [treenode, loadhide, right = of mainagent] (inner0) {};
		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner0] (inner1) {};
		\node [hidenode, below right = 0.5cm and 1cm of inner0] (inner2) {};
		\node [treenode, loadhide, below right = 0.5cm and 1cm of inner1] (inner3) {};
		\node [treenode, loadhide, right = of inner3] (inner4) {};
		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
		\node [hidenode, below right = 0.5cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};

@@

\subsection{Action Space}
When monitoring a source-destination pair, an agent uses its state vector to decide which proportion of that flow's \emph{inbound} traffic should be dropped.
%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
We choose to drop packets rather than impose traffic limits as it affords us a discrete action space without prior knowledge of traffic characteristics or measurement.
Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
We offer two models on how to choose $p$:

@@

\subsubsection{Guarded control}
The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$, choosing to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.

The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.
To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
Most importantly, an agent must choose to punish a flow multiple times in succession to cause rapid degradation, reducing variance while allowing an agent to see how a host reacts to structured changes in the environment.

As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic.

\subsubsection{Risks}
Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate UDP traffic is proportional to $1 - p$.
%It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which depends on $1/p^{0.75}$ \cite{rfc8312}.
This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
%As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.

@@

\subsection{Systems Considerations}
Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
This introduces some potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
To adapt to these problems, we make use of \emph{timed random sequential} updates.

@@

\fakepara{Global state}
These are the vector of load measurements along a flow's path introduced in \cref{sec:feature-space}.
These indicate the overall health of the network, and crucially are all measurements which an agent directly controls.

@@

\fakepara{Last action taken}
This encodes an agent's current belief in the maliciousness of a flow.
This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
It's important to note that this feature only makes sense once combined with another flow feature.

@@

All of the above features, save for global state, are 1-dimensional.
\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same techniques is applied with UDP traffic (omitted).
%?? Core findings---different protocols need different features, so everything we proposed above has a use!
The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
All features converge to their highest-observed performance within around \num{4000} timesteps.
In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
%?? How do they do when combined after individual training? Pretty well, especially for TCP.
%Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
%In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.

@@

%From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
%As it turns out, 
%Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
We contribute network models built around live testing of reactive TCP and UDP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving their real-time dynamics in a way that trace-based evaluation cannot.
First and foremost, we are interested in representative load and packet inter-arrival characteristics and in how these characteristics evolve in response to actions.
We introduce these models because are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
%Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
%If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.

@@

\subsection{Network Design}
We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
%We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.

\subsection{TCP (HTTP) Traffic Model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate TCP traffic, the server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
Hosts and clients use the TCP Cubic \cite{rfc8312} flavour of TCP.
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.

>>>>

\begin{abstract}
%?? Trim to 75--200 words, simplify.
DDoS attacks plague the availability of online services today, yet like many cybersecurity problems are evolving and non-stationary.
Normal and attack patterns shift as new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
Accordingly, it is difficult to apply machine learning-based techniques and defences in practice.
%Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.

@@

Our results show the existence of flow features with high predictive power for different traffic classes, when used as a basis for feedback-loop-like control.
We show that the new RL agent models can offer a significant increase in goodput of legitimate TCP traffic for many choices of host density.

@@

To date, there have been few applications of this class of algorithms towards intrusion detection and prevention which make use of their full potential for online control, rather than using them as the basis for a classifier.
We aim to take steps to redress this and establish their proper capabilities, beyond simple ``blind application''.
%?? Expand as required
What approaches do exist are aimed towards the task of adaptive online DDoS mitigation, and rely upon learning to control probabilistic packet drop.
%?? THAT IS A MAJOR CONTRIB, MENTION IT EVERYWHERE YOU CAN

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An empirical evaluation of our new models against the state-of-the-art in RL-based DDoS mitigation using these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
\end{itemize}

@@

\emph{Distributed denial of service} (DDoS) attacks are concentrated efforts by many hosts to reduce the availability of a service, typically to inflict financial harm or as an act of vandalism.
Attackers achieve this by either exploiting peculiarities of operating system or application behaviour in \emph{semantic attacks} (e.g., \emph{SYN flooding attacks}), or overwhelming their target through sheer volume of requests or inbound packets (\emph{volume-based attacks}) \cite{DBLP:conf/imc/JonkerKKRSD17}.
Hosts often participate unwillingly, typically having been recruited into a \emph{botnet} by malware infection to be orchestrated from elsewhere \cite{DBLP:conf/uss/AntonakakisABBB17}.

@@

\begin{figure}
	\centering
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
		\node at (0,0.3){$s = \begin{pmatrix}
			0.7 \\
			0.3
			\end{pmatrix}$};
		\node at (0,-1) {$\bm{x}(s,\cdot) = \begin{Bmatrix}
			T_{1,9}, \\
			T_{2,5}, \\
			T_{\mathit{bias}}
			\end{Bmatrix}$};

@@

			\draw[step=0.5cm,color=uofgpumpkin,opacity=0.9,shift={(0.25,-0.25)},label=above:{Tiling 1}] (-0.5,-0.5) grid (1,1);
			\fill[uofgpumpkin,opacity=0.5,shift={(0.25,-0.25)}] (0,0) rectangle (0.5,0.5);
			\node[color=uofgpumpkin!50!uofgrust] at (0.25,-0.95) {\footnotesize Tiling 2};

@@

		\end{tikzpicture}
		}
		\caption{Tile Coding\label{fig:tilecode-code}}
	\end{subfigure}%
	\begin{subfigure}{0.45\linewidth}
		\resizebox{\linewidth}{!}{
		\begin{tikzpicture}
		% Top half
		\def\topacs{
			{-0.3,-0.5,-0.1,0.8},
			{0.1,0.1,-0.2,0.3},
			{0.3,0.4,0.2,-0.4}%
		}

@@

This combination of algorithm and coding strategy is well-optimised, if actions are discrete; this allows a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each tile.
Action values for any state are then obtained by summing the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
This means that we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update:
\begin{equation}
\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
\label{eqn:sg-sarsa-opt}
\end{equation}
\Cref{fig:tilecode-update} shows how this applies to our prior example.
If desired we may define a state space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.
A caveat of tile coding remains, in that the value of $\alpha$ must be reduced according to the number of tilings to prevent divergence at the expense of slower learning ($\alpha \leftarrow \alpha / n_{\mathit{tilings}}$).

@@

\subsubsection{Action rate}
We chose to adapt the algorithm to prioritise rapid response to changes in network state and to visit as many state-to-state transitions as possible to facilitate effective learning.
To this end, we allow agents to make many decisions per timestep.
We maintain the last state-action pair associated with each source IP and destination, and calculate any actions for the flows which still exist.
Finally, we update $\wvec{}$ using each available trace and the reward signal from the relevant destination.
As exploration still occurs for each action, this approach reduces $\epsilon$ multiple notches every timestep.
In turn, we increase the annealing window for $\epsilon$ by a factor of \num{2.67} so as to preserve exploration over time, by accounting for the greater volume of decisions being made.

@@

\begin{figure}
	\centering
	\resizebox{0.8\linewidth}{!}{
		\begin{tikzpicture}[
		texts/.style = {text=black},
		labeltexts/.style = {text=gray},
		treeline/.style = {draw=uofgburgundy},
		treenode/.style = {texts, circle, centered, fill=white, treeline},
		load/.style = {fill=uofgcobalt},
		loadline/.style = {draw=uofgcobalt, line width=0.75mm},
		loadhide/.style = {fill=uofgcobalt!40!white},
		external/.style = {fill=uofgrust},
		externalhide/.style = {fill=uofgrust!40!white},
		hideline/.style = {draw=uofgsandstone!40!white},
		hidenode/.style = {treenode, hideline},
		grow'=right
		]
		\node [treenode, loadhide, label={[texts]above:Agent 1}] (mainagent) {};
		\node [treenode, loadhide, right = of mainagent] (inner0) {};
		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner0] (inner1) {};
		\node [hidenode, below right = 0.5cm and 1cm of inner0] (inner2) {};
		\node [treenode, loadhide, below right = 0.5cm and 1cm of inner1] (inner3) {};
		\node [treenode, loadhide, right = of inner3] (inner4) {};
		\node [treenode, loadhide, above right = 0.5cm and 1cm of inner4, label={[texts]above:$s_0$}] (s0) {};
		\node [hidenode, below right = 0.5cm and 1cm of inner4, label={[labeltexts]above:$s_1$}] (s1) {};

@@

\subsection{Action Space}
When monitoring a source-destination pair, an agent uses its state vector to decide which proportion of that flow's \emph{inbound} traffic should be dropped.
%?? Why pdrop? Allows for discrete action space, don't have to account for buckets/fairness/burstiness.
This is implemented by installing an action via OpenFlow, instructing its host switch to drop each relevant packet with probability $p$.
We choose to drop packets rather than impose traffic limits as it offers us a discrete action space without prior knowledge of traffic characteristics or measurement.
Furthermore, we need not consider burstiness, fairness or implementation costs (such as per-flow bucket sizes) which could limit scalability.
We offer two models on how to choose $p$:

@@

\subsubsection{Guarded control}
The measurements of \textcite{DBLP:conf/ndss/KangGS16} suggest that bot attack flows cannot scale up to match an increase in available bandwidth.
We apply their observations within the RL paradigm by constraining how an agent treats each flow using a simple finite state automaton: we restrict $p \in \left\{ 0.00, 0.05, 0.25, 0.50, 1.0 \right\}$,.
The action set is then simply to \emph{maintain}, \emph{increase}, or \emph{decrease} $p$ for a flow in single steps.

We choose these potential values for $p$ to add complete filtering to a steady progression of rate-limiters (\SI{25}{\percent} increments for UDP traffic).
The outlier, $p=0.05$, corresponds to roughly a \SI{50}{\percent} rate reduction for TCP flows in our test topology.
This uneven spread of choices for $p$ allows light and heavy rate reduction to be applied to both congestion-aware and non-congestion-aware traffic as required.

To enable temporary bandwidth expansion in all deployments, every flow is initially placed under light packet drop ($p=0.05$); this is chosen above the equivalent for UDP due to TCP's higher prevalence.
Most importantly, an agent must now choose to punish a flow multiple times in succession to cause rapid degradation, reducing variance while allowing an agent to see how a host reacts to structured changes in the environment.

As each agent now requires the capability to plan ahead, we require a discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing.

\subsubsection{Risks}
Our mode of action means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
This is particularly severe when handling TCP traffic: the Mathis equation \cite{DBLP:journals/ccr/MathisSMO97} states that TCP bandwidth is proportional to $1/\sqrt{p}$ (noting that $p$ is nonzero in any real network) while constant bitrate (CBR) UDP traffic is proportional to $1 - p$.
%It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.
This weakness still presents itself in modern flavours of TCP, such as TCP Cubic which in turn has bandwidth proportional to $1/p^{0.75}$ \cite{rfc8312}.
This is of particular importance due to the prevalence of TCP and other congestion-aware protocols within the Internet.
Our own analysis of CAIDA datasets \cite{caida-2018-passive} shows that congestion-aware traffic makes up at least \SIrange{73}{82}{\percent} of packets, corresponding to \SIrange{77}{84}{\percent} of data volume\footnote{\url{https://github.com/FelixMcFelix/caida-stats}}.
QUIC, a future congestion-aware protocol, comprises \SIrange{2.6}{9.1}{\percent} of traffic observed on backbone links, depending on location and typical workload \cite{DBLP:conf/pam/RuthPDH18}.
%As far as future network protocols are concerned, QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}, a congestion-aware stream transmission protocol, will behave much like TCP, showing the importance of further development to properly handle traffic with such characteristics.

@@

\subsection{Systems Considerations}
Taking many actions per-timestep means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
This introduces some potential issues: the inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated.
At their worst, these risks present additional attack surface to an adversary.
To adapt to these problems, we make use of \emph{timed random sequential} updates.

@@

\fakepara{Global state}
This is the vector of load measurements along a flow's path introduced in \cref{sec:feature-space}.
These values indicate the overall health of the network, and crucially are all measurements which an agent directly controls.

@@

\fakepara{Last action taken}
This encodes an agent's current belief in the maliciousness of a flow.
This feature also potentially allows forgiveness, serving as a reference point for determining whether a source mistakenly marked as malicious exhibits different falloff behaviour after punishment.
It's important to note that this feature only makes sense once combined with another flow feature, and never appears individually tile-coded.

@@

All of the above features, save for global state, are 1-dimensional.
\Cref{fig:udp-feature-plots} shows the effectiveness of each feature for UDP (resp.\ \cref{fig:tcp-cap-feature-plots} for TCP), using a single-destination topology (\cref{sec:single-dest}) with $n=2$ hosts per egress point averaged over 10 episodes.
\Cref{fig:tcp-laf-feature-plots} demonstrates how feature accuracy varies when tiled alongside \emph{last action}, with similar trends being observed when the same technique is applied with UDP traffic (omitted).
%?? Core findings---different protocols need different features, so everything we proposed above has a use!
The plots demonstrate that different protocols and traffic classes are best defended by different features---as such, every feature presented has value in a complete model.
All features converge to their highest-observed performance within around \num{4000} timesteps.
In general, some of the most effective features are the global state, mean IAT, mean inbound packet size and $\Delta$ rates.
%?? How do they do when combined after individual training? Pretty well, especially for TCP.
%Additional testing shows that the learned per-feature policies may be easily combined (by summing action values), and that this technique is particularly effective for TCP; these results are omitted to preserve space.
%In no cases, however, do we manage to completely block attack traffic---at convergence, we observe that system load remains consistently at $U_s$.

@@

%From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
%As it turns out, 
%Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
We contribute network models built around live testing of reactive TCP and UDP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving their real-time dynamics in a way that trace-based evaluation cannot.
First and foremost, we are interested in representative load and packet inter-arrival characteristics and in how these characteristics evolve in response to actions.
We introduce these models because we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
%Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
%If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.

@@

\subsection{Network Design}
We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
If several ports offer different (equal-length) paths to a destination, a consistent random port is chosen from the flow-hash by an OpenFlow \emph{Group action} (in \emph{select} mode).
If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
The controller is also responsible for computing how switches respond to ARP requests: this need arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
%We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.

\subsection{TCP (HTTP) Traffic Model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate TCP traffic, server nodes run an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
Hosts and clients use the TCP Cubic \cite{rfc8312} flavour of TCP.
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it, and requests several random files known to exist within a website, followed by any dependent resources for each (stylesheets, images, etc.).
On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime.
This presents a more balanced distribution of flow duration and size, with large files included to model elephant flows.
